### 2017-07-31 Tran Nott Kohn/outline.tex
PF
Particle Filter in good condition
Various updates

@@

In this case, using $f(\boldsymbol{\epsilon}, \boldsymbol{\lambda}) = \{\boldsymbol{\theta}, x_{T+1}\}$,
\begin{equation}
\label{DimRedLogJoint}
\ln(p(y_{1:T}, f(\boldsymbol{\epsilon}, \boldsymbol{\lambda}))) = \ln\left(\int_{X_T} p(x_{T+1} | \theta, x_T)p(x_T | y_{1:T}, \boldsymbol{\theta})d_{X_T}\right) + \ln(p(y_{1:T} |\boldsymbol{\theta})) + \ln(p(\boldsymbol{\theta}))
\end{equation}
In many cases, the distributions $p(x_T | y_{1:T}, \boldsymbol{\theta})$ and $p(y_{1:T} | \boldsymbol{\theta})$ are intractable, but \citet{TranM2017} note that the particle filter estimators $\hat{p}(x_T | y_{1:T}, \boldsymbol{\theta})$ and $\hat{p}(y_{1:T} | \boldsymbol{\theta})$ are sufficient substitutes in what they refer to as Variational Bayes with Intractable Likelihood (VBIL). The particle filter estimation of the distribution $\hat{p}(x_T | y_{1:T}, \boldsymbol{\theta})$ is expressed as a discrete set of point masses $x_T^{(k)}$ and weights $\pi_T^{(k)}$, where $k = 1, \dots, N$ represent the particles of the particle filter. In this case, the integral in (\ref{DimRedLogJoint}) reduces to the sum
\begin{equation}
\label{ParticleFilterIntegral}
\int_{X_T} p(x_{T+1} | \theta, x_T)p(x_T | y_{1:T}, \boldsymbol{\theta})d_{X_T} = \sum_{k=1}^N \pi_T^{(k)} p(x_{T+1} | \boldsymbol{\theta}, x_T^{(k)}),
\end{equation}
and hence
\begin{equation}
\label{PFLogJoint}
\ln(p(y_{1:T}, f(\boldsymbol{\epsilon}, \boldsymbol{\lambda}))) \approx \ln \left(\sum_{k=1}^N \pi_T^{(k)} p(x_{T+1} | \boldsymbol{\theta}, x_T^{(k)})\right) + \ln(\hat{p}(y_{1:T} |\boldsymbol{\theta})) + \ln(p(\boldsymbol{\theta}))
\end{equation}
can be substituted into (\ref{ReparamDeriv}). This derivative of (\ref{PFLogJoint}) with respect to $\boldsymbol{\lambda}$ can be obtained with automatic differentiation tools such as those provided by Stan \citep{Carpenter2015}.

@@

$$p(y_{1:T+S}, \theta, x_T, x_{T+S}) = p(\theta) p(y_{1:T+S} | \theta) p (x_T | \theta, y_{1:T+S}) p(x_{T+S} | x_T, \theta, y_{1:T+S})$$

\begin{itemize}
\item Proper updates with filter/smooth
\item Secondary approximation - no smoothing
\item Straight up particle filter on old approximation
\item PF on old MCMC
\item Simulation Results
\end{itemize}

>>>>

\citep{Arulampalam2002}

@@

In this case, using $f(\boldsymbol{\epsilon}, \boldsymbol{\lambda}) = \{\boldsymbol{\theta}, x_{T+1}\}$,
\begin{equation}
\label{DimRedLogJoint}
\ln(p(y_{1:T}, f(\boldsymbol{\epsilon}, \boldsymbol{\lambda}))) = \ln\left(\int_{X_T} p(x_{T+1} | \theta, x_T)p(x_T | y_{1:T}, \boldsymbol{\theta})d_{X_T}\right) + \ln(p(y_{1:T} |\boldsymbol{\theta})) + \ln(p(\boldsymbol{\theta}))
\end{equation}
In many cases, the distributions $p(x_T | y_{1:T}, \boldsymbol{\theta})$ and $p(y_{1:T} | \boldsymbol{\theta})$ are intractable, but \citet{TranM2017} note that the particle filter estimators $\hat{p}(x_T | y_{1:T}, \boldsymbol{\theta})$ and $\hat{p}(y_{1:T} | \boldsymbol{\theta})$ are sufficient substitutes in what they refer to as Variational Bayes with Intractable Likelihood (VBIL). The particle filter estimation of the distribution $\hat{p}(x_T | y_{1:T}, \boldsymbol{\theta})$ is expressed as a discrete set of point masses $x_T^{(k)}$ and weights $\pi_T^{(k)}$, where $k = 1, \dots, N$ represent the particles of the particle filter. In this case, the integral in (\ref{DimRedLogJoint}) reduces to the sum
\begin{equation}
\label{ParticleFilterIntegral}
\int_{x_T} p(x_{T+1} | \theta, x_T)p(x_T | y_{1:T}, \boldsymbol{\theta})d_{x_T} = \sum_{k=1}^N \pi_T^{(k)} p(x_{T+1} | \boldsymbol{\theta}, x_T^{(k)}),
\end{equation}
and hence
\begin{equation}
\label{PFLogJoint}
\ln(p(y_{1:T}, f(\boldsymbol{\epsilon}, \boldsymbol{\lambda}))) \approx \ln \left(\sum_{k=1}^N \pi_T^{(k)} p(x_{T+1} | \boldsymbol{\theta}, x_T^{(k)})\right) + \ln(\hat{p}(y_{1:T} |\boldsymbol{\theta})) + \ln(p(\boldsymbol{\theta}))
\end{equation}
can be substituted into (\ref{ReparamDeriv}). This derivative of (\ref{PFLogJoint}) with respect to $\boldsymbol{\lambda}$ can be obtained with automatic differentiation tools such as those provided by Stan \citep{Carpenter2015}.

@@

\begin{itemize}
\item We have an approximation of $q(\theta, x_{T+1} | y_{1:T})$ which is a great start. But if we observe data up to time $T+S$ I'm interested in incorporating this to update the approximation to $q(\theta, x_{T+S+1} | y_{1:T+S})$.
\item Note: The approxmation going up to state $x_T, x_{T+1}, x_{T+S}, x_{T+S+H}$ doesn't really matter. An approximation at one point in time allows us to obtain an approximation for a different time due to the linear/gaussian transition equation. The original approximation could easily be modified to $q(\theta, x_{T+S+1} | y_{1:T})$, but what I think is important is conditioning this on $y_{T+1:T+S}$.
\item The form of the joint distribution, $p(y_{1:T+S}, \theta, x_{T+S+1})$ should follow the same decomposition as before,
$$p(y_{1:T+S}, \theta, x_{T+S+1}) = p(\theta) p(y_{1:T+S} | \theta) \int_{X_{T+S}} p(X_{T+S+1} | \theta, x_{T+S})p(x_{T+S} | \theta, y_{1:T+S})dX_{T+S},$$
using a particle filter to estimate the intractable distributions.
\item Key step is using information in the approximation of $q(\theta, x_T | y_{1:T})$ to help form $q(\theta, x_{T+S} | y_{1:T+S})$.
\item We have access to the approximation of the conditional distribution, $q(x_T | \theta, y_{1:T})$ by integration of VB fit.
\item Start the Particle Filter with $x_T$, using VB approximation $q(x_T | \theta, y_{1:T})$ as a proposal distribution for $x_T$, instead of the prior as is commonly used.
\item The importance sampling weights should be easily obtainable using this new proposal.
\item Then the filter needs to be run forward for $S$ steps.
\item Using a VB approximation as an importance sampling proposal distribution is a new contribution. Hopefully it'll work well.
\end{itemize}
